---
output:
  pdf_document: default
  html_document: default
---
## Uber Insights: A Ride Through Data
### Load necessary libraries
```{r}
library(ggplot2)
library(dplyr)
library(lubridate)
```


### Read the dataset
```{r}
data <- read.csv("uber-raw-data-aug14.csv")
head(data)
```
### Examine the structure of the Date.Time column to understand its format and then convert Date.Time from character format to POSIXct for easier date-time manipulation.
### The next step is to extract additional date-time features: Day, Weekday, and Hour

```{r}
str(data$Date.Time)

# Convert Date.Time from character to POSIXct
data$Date.Time <- as.POSIXct(data$Date.Time, format = "%m/%d/%Y %H:%M:%S")

data$Day <- day(data$Date.Time)
data$Weekday <- wday(data$Date.Time, label = TRUE)
data$Hour <- hour(data$Date.Time)

```

### Create a bar plot to visualize the number of Uber trips per day
```{r}
ggplot(data, aes(x = Day)) +
  geom_bar(fill = "blue") +
  labs(title = "Daily Uber Trips", x = "Day", y = "Number of Trips") +
  theme_dark()

```
### Create a bar plot to visualize the number of Uber trips per hour
```{r}
ggplot(data, aes(x=Hour)) + 
  geom_bar(fill = "brown") + 
  labs(title = "Hourly Uber Trips", x = "Hours", y = "Number of Trips") + 
  theme_dark()
```
### Create a bar plot to visualize the number of Uber trips by weekday

```{r}
ggplot(data, aes(x=Weekday)) + 
  geom_bar(fill = "lightyellow")+
  labs(title = "Weekly Uber Trips", x = "Days of the Week", y = "Number of trips") +
  theme_dark()
```
### Display summary statistics of the dataset
```{r}
summary(data)
```
### Count the number of trips grouped by Weekday and Hour
```{r}
# Count trips by weekday and hour
trip_counts <- data %>%
  group_by(Weekday, Hour) %>%
  summarise(Trip_Count = n())

# View the first few rows of the summary
head(trip_counts)

```
### Create a heatmap to visualize the number of Uber trips by Hour and Weekday
```{r}
# Create a heatmap
ggplot(trip_counts, aes(x = Hour, y = Weekday, fill = Trip_Count)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "white", high = "blue") +
  labs(title = "Heatmap of Uber Trips by Hour and Weekday", x = "Hour", y = "Weekday") +
  theme_dark()
```
### Count the number of trips by Latitude and Longitude to visualize geographic distribution and create a scatterplot to show the number of Uber trips based on Latitude and Longitude
```{r}
trip_counts_geo <- data %>%
  group_by(Lat, Lon) %>%
  summarise(Trip_Count = n(), .groups = 'drop')

ggplot(trip_counts_geo, aes(x = Lon, y = Lat, size = Trip_Count, color = Trip_Count)) +
  geom_point(alpha = 0.4) +
  scale_color_gradient(low = "white", high = "blue") +
  labs(title = "Number of Uber Trips by Latitude and Longitude",
       x = "Longitude",
       y = "Latitude",
       size = "Number of Trips") +
  theme_dark()
```




## Summary of Uber Trip Analysis

### The analysis of Uber trip data for August 2014 reveals several key insights:

### Peak Profitability: Friday stands out as the most profitable day for Uber, indicating a higher demand for rides as the weekend approaches.
### Lower Usage on Mondays: Monday shows the least number of trips, suggesting that fewer people opt for Uber at the beginning of the week.
### Busiest Hour: The period between 5 PM and 6 PM is identified as the busiest hour for Uber, likely due to the end of the workday when many commuters seek rides.
### Early Morning Demand: There is a noticeable increase in Uber trips starting around 6 AM, indicating that early morning rides are common, possibly for commuters heading to work.
### Geographic Concentration: Most Uber trips originate in the Manhattan region of New York, highlighting this area as a central hub for ride requests.
### These insights can be valuable for understanding rider behavior and optimizing service strategies for Uber.
